---
title: "Methods"
---

```{r}
#| label: setup
#| include: false
source("_common.R")
```

## Data Source and Study Population

The analysis utilized the 2022 Bangladesh Demographic and Health Survey (BDHS), a nationally representative cross-sectional survey [@niport2024bdhs]. The study population was restricted to ever-married women aged 15–49 who had a live birth in the three years preceding the survey to ensure the availability of maternal healthcare and decision-making indicators.

```{mermaid}
graph TD
    A["Ever-married women interviewed (15–49 years)<br/>(N = 30,344)"]

    E1["Excluded: No live birth in the 3 years<br/>preceding the survey (n = 18,687)"]

    B["<b>Analytic sample for descriptive statistics</b><br/>(n = 11,657)"]

    E2["Excluded: Missing values in multivariable<br/>model covariates (n = 4,031)"]

    C["<b>Final sample for multivariable modeling</b><br/>(n = 7,626)"]

    A --> B
    A -.-> E1
    B --> C
    B -.-> E2

    %% Styling
    style B fill:#f0f7ff,stroke:#0056b3,stroke-width:2px
    style C fill:#f4fff4,stroke:#2e7d32,stroke-width:2px
```

Due to the nature of survey skip patterns and item-level non-response, a complete-case approach requiring all nine risk indicators to be observed would have restricted the sample to 1,531 women (13.1% of those with recent births). To maintain population representativeness and statistical power, the primary analysis utilized an **available-indicator** cumulative score, including all 11,657 women for descriptive analysis and 7,626 women for the multivariable modeling where covariate data were complete.

## Outcome and Risk Domains

Nine binary risk indicators were grouped into three functional domains to capture the multi-dimensional nature of the syndemic burden.

```{r}
#| label: risk-domain-table
tibble(
  Domain = c("Nutritional", "Reproductive health", "Healthcare access and autonomy", "**Total Syndemic Score**"),
  Indicators = c(
    "Underweight (BMI < 18.5); stunted height (< 145 cm)",
    "Adolescent first birth (< 20y); high parity (> 4); short birth interval (< 24m); home delivery",
    "Inadequate ANC (< 4 visits); distance to facility (major problem); low health autonomy",
    "Sum of all nine binary indicators (dichotomized 0/1)"
  ),
  `Score Range` = c("0–2", "0–4", "0–3", "**0–9**")
) |>
  kable(caption = "Syndemic risk domains, indicators, and score construction", align = "llc", escape = FALSE)
```

For descriptive reporting and sensitivity analysis, the total score was collapsed into three ordered categories: **Low burden** (score 0), **Medium burden** (score 1–3), and **High burden** (score ≥ 4).

## Covariates and Model Parameters

Multivariable models included sociodemographic factors identified in the literature as key drivers of maternal and nutritional health in Bangladesh.

```{r}
#| label: covariate-table
tibble(
  `Variable Group` = c("Demographic", "Socioeconomic", "Geography", "Empowerment"),
  `Covariates (Levels)` = c(
    "Age group (15–24, 25–34, 35–49)",
    "Respondent education (No education, Primary, Secondary, Higher); Wealth quintile (Poorest to Richest); Husband's education",
    "Administrative Division (8 regions); Type of residence (Urban, Rural)",
    "Employment status (Working, Not working)"
  ),
  `Reference Category` = c("15–24 years", "Higher education; Richest quintile", "Dhaka Division; Urban residence", "Not working")
) |>
  kable(caption = "Covariates included in multivariable regression models", align = "lll")
```

## Statistical Analysis

All descriptive and regression analyses incorporated BDHS survey weights, strata, and primary sampling units using R's `survey` package [@lumley2004survey; @lumley2010complex].

**Primary model:** Survey-weighted quasi-Poisson regression (`svyglm`, `family = quasipoisson`) for the raw syndemic score (0–8). The quasi-Poisson family was selected because the score is a bounded integer with mild overdispersion relative to a simple Poisson process (observed variance/mean ratio = 1.23; svyglm dispersion φ = 0.77; diagnostic weighted GLM φ = 1.08). Negative binomial regression was evaluated but yielded an equivalent fit (θ = 22.20), confirming that quasi-Poisson was sufficient. Results are reported as incidence rate ratios (IRR) with 95% confidence intervals [@hilbe2011negbin; @gardner1995regression].

**Sensitivity analysis:** Ordinal proportional-odds logistic regression for low/medium/high burden categories [@mccullagh1980regression; @liu2016proportional]. The ordinal model provides a complementary perspective using the collapsed categorical outcome; its results are reported in the [Supplementary Appendix](appendix.qmd).

## References

::: {#refs}
:::
